---
title: "Statistical Computing Homework 6"
# subtitle: "possible subtitle goes here"
author:
  - Jieying Jiao
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
papersize: letter
fontsize: 11pt
bibliography: template.bib
biblio-style: asa
keywords: Template, R Markdown, bookdown, Data Lab
# keywords set in YAML header here only go to the properties of the PDF output
# the keywords that appear in PDF output are set in latex/before_body.tex
output:
  bookdown::pdf_document2
abstract: |
    This is Jieying Jiao's homework 6 for statistical computing, fall 2018.
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## some utility functions, see the source code for details
source("utils_template.R")

## specify the packages needed
pkgs <- "ggplot2"
need.packages(pkgs)

## external data can be read in by regular functions,
## such as read.table or load

## for latex and html output
isHtml <- knitr::is_html_output()
isLatex <- knitr::is_latex_output()
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```

# Exercise 5.1.1

## Normalize g

\begin{align*}
&\int_0^\infty (2x^{\theta-1}-x^{\theta - 1/2}) e^{-x} dx\\
&= 2\int_0^\infty x^{\theta-1}e^{-x} dx + \int_0^\infty x^{\theta+1/2-1}e^{-x}dx\\
&= 2\Gamma(\theta) \frac{1}{\Gamma(\theta)}\int_0^\infty x^{\theta-1}e^{-x}dx + 
\frac{\Gamma(\theta+1/2)}{\Gamma(\theta+1/2)}\int_0^\infty x^{\theta+1/2-1}e^{-x}dx\\
&= 2\Gamma(\theta)+\Gamma(\theta+1/2)\\
\\
\Rightarrow C &= \frac{1}{2\Gamma(\theta)+\Gamma(\theta+1/2)}\\
\\
\Rightarrow g(x) &= \frac{2\Gamma(\theta)}{2\Gamma(\theta)+\Gamma(\theta+1/2)} 
\frac{1}{\Gamma(\theta)}x^{\theta-1}e^{-x} + \frac{\Gamma(\theta+1/2)}{2\Gamma(\theta)+
\Gamma(\theta+1/2)}\frac{1}{\Gamma(\theta+1/2)}x^{\theta+1/2-1}e^{-x}\\
&= \frac{2\Gamma(\theta)}{2\Gamma(\theta)+\Gamma(\theta+1/2)} Gamma(\theta, 1) +
\frac{\Gamma(\theta+1/2)}{2\Gamma(\theta)+\Gamma(\theta+1/2)} Gamma(\theta+1/2, 1)\\
\end{align*}

So g is a mixture of $Gamma(\theta, 1)$ and $Gamma(\theta+1/2, 1)$, with weights 
shown in the above formula.

## Draw sample from g, mixture gamma

Pesudo-Code:

(1)Step1: Sample $U \sim Unif(0, 1)$

(2)Step2: 

if $U < \frac{2\Gamma(\theta)}{2\Gamma(\theta)+\Gamma(\theta+1/2)}${

sample $X \sim Gamma(\theta, 1)$

}

else {sample $X \sim Gamma(\theta+1/2, 1)$}

(3)Step3: Return X

```{r HW6_1, warning = FALSE}
rmixgamma <- function(shape, scale, probs, n) {
  x <- rep(0, n)
  for (i in 1:n) {
    u <- runif(1, 0, 1)
    if(u < probs[1]) {
      x[i] <- rgamma(1, shape = shape[1], scale = scale[1])
    } else x[i] <- rgamma(1, shape = shape[2], scale = scale[2])
  }
  return(x)
}

theta <- 1
shape <- c(theta, theta+1)
scale <- c(1, 1)
probs <- c(2*gamma(theta)/(2*gamma(theta)+gamma(theta+1/2)), 
           gamma(theta+1/2)/(2*gamma(theta)+gamma(theta+1/2)))
n <- 10000
X <- rmixgamma(shape = shape, scale = scale, probs = probs, n = n)
hist(X, nclass = 20, probability = TRUE, ylim = c(0, 1))
points(density(X), type = "l")
gfunc <- function(x, para) {
  y <- (2*x^(para-1)+x^(para-0.5))*exp(-x)/(2*gamma(para)+gamma(para+0.5))
  return(y)
}
curve(gfunc(x, para = theta), from = 0, to = 12, add = TRUE, col = "red")
```

## Sample from f, use rejection sampling

\begin{align*}
q(x) &= \sqrt{4+x}x^{\theta-1}e^{-x}\\
g(x) &= C(2x^{\theta-1}+x^{\theta-1/2})e^{-x}\\
\Rightarrow \alpha &= sup_{x>0}\frac{q(x)}{g(x)} = \frac{1}{C}sup_{x>0}\frac{\sqrt{4+x}}{2+x^{1/2}}
 = \frac{1}{C}\\
\Rightarrow \alpha g(x) &= (2x^{\theta-1}+x^{\theta-1/2})e^{-x}\\
\end{align*}

Pesudo-Code:

(1)Step1: Sample $X\sim g$, $U\sim Unif(0, 1)$.

(2)Step2: if $U > \frac{q(x)}{\alpha g(x)}$ {

then go to step 1

} else return X

(3)Step3: repeat step 1-2, until get desired number of sample.

```{r HW6_2, warning=FALSE}
rffunc <- function(n) {
  y <- rep(0, n)
  count <- 0
  while (count < n) {
    u <- runif(1, 0, 1)
    x <- rmixgamma(shape = shape, scale = scale, probs = probs, n = 1)
    if(u <= sqrt(4+x)/(2+sqrt(x))) {
      count <- count + 1
      y[count] <- x
    }
  }
  return(y)
}

X <- rffunc(10000)
hist(X, nclass = 20, probability = TRUE, ylim = c(0, 1))
points(density(X), type = "l", col = "red")
```

# Exercise 5.1.2

## Use mixture Beta distribution
I design g to be the mixture of $Beta(\theta, 1)$ and $Beta(1, \beta)$, with equal 
weights:

\begin{align*}
g(x) &= \frac{1}{2Beta(\theta, 1)}x^{\theta-1}+\frac{1}{2Beta(1, \beta)}(1-x)^{\beta-1}\\
     &= C_1x^{\theta-1}+C_2(1-x)^{\beta-1}\\
\Rightarrow \alpha &= sup_{x\in(0, 1)} \frac{\frac{x^{\theta-1}}{1+x^2}+
\sqrt{2+x^2}(1-x)^{\beta-1}}{c_1x^{\theta-1}+C_2(1-x)^{\beta-1}}\\
                   &= sup_{x\in(0, 1)}\frac{x^{\theta-1}+\sqrt{2+x^2}(1+x^2)(1-x)^{\beta-1}}
                   {C_1x^{\theta-1}(1+x^2)+C_2(1-x)^{\beta-1}(1+x^2)}\\
\end{align*}

Easy to observe that the above fraction is bounded on $x\in(0, 1)$, so $\alpha$ 
exists. For simplicity of calculation, we make $\theta = 2, \beta = 1$, then:
\begin{align*}
C_1 &= \frac{1}{2Beta(\theta, 1)}=1\\
C_2 &= \frac{1}{2Beta(1, \beta)}=\frac{1}{2}\\
\alpha &= sup_{x\in(0, 1)}\frac{x+\sqrt{2+x^2}(1+x^2)}{x(1+x^2)+\frac{1}{2}(1+x^2)}=2\sqrt{2}\\
\alpha g(x) &= 2\sqrt{2}x+\sqrt{2}\\
\end{align*}

Pesudo-Code:

(1)Step1: Sample $X\sim g$, $U\sim Unif(0, 1)$.

(2)Step2: if $U > \frac{q(x)}{\alpha g(x)}$ {

then go to step 1

} else return X

(3)Step3: repeat step 1-2, until get desired number of sample.

```{r HW6_3, warning=FALSE}
rffunc2 <- function(n) {
  y <- rep(0, n)
  count <- 0
  while (count < n) {
    u <- runif(1, 0, 1)
    if (u < 0.5) {
      x <- rbeta(1, shape1 = 2, shape2 = 1)
    } else x <- rbeta(1, shape1 = 1, shape2 = 1)
    if (u <= (x/(1+x^2)+sqrt(2+x^2))/sqrt(2)/(2*x+1)) {
      count <- count+1
      y[count] <- x
    }
  }
  return(y)
}
X <- rffunc2(10000)
hist(X, nclass = 20, probability = TRUE)
points(density(X), type = "l", col = "red")
```


## Use seperate Beta distribution

Let $g_1(x) \sim Beta(\theta, 1)$, $g_2(x) \sim Beta(1, \beta)$:
\begin{align*}
g_1(x) &= \frac{1}{Beta(\theta, 1)}x^{\theta-1}\\
g_2(x) &= \frac{1}{Beta(1, \beta)}(1-x)^{\beta-1}\\
\alpha_1 &= sup_{(0, 1)}\frac{q_1(x)}{g_1(x)} = sup_{(0, 1)}\frac{Beta(\theta, 1)}
{1+x^2}=Beta(\theta, 1)\\
\alpha_2 &= sup_{(0, 1)}\frac{q_2(x)}{g_2(x)} = sum_{(0, 1)}Beta(1, \beta)\sqrt{2+x^2}=
\sqrt{3}Beta(1, \beta)\\
p_1 &= \frac{\alpha_1}{\alpha_1+\alpha_2}\\
p_2 &= \frac{\alpha_2}{\alpha_1+\alpha_2}\\
\end{align*}

Pesudo-Code:

(1)Step1: Sample k from $\{1, 2\}$ with probs $p_1, p_2$

(2)Step2: Sample $X \sim g_k$, $U \sim Unif(0, 1)$

(3)Step3: if $U > \frac{q_k(X)}{\alpha_kg_k(X)}$ {

then go to step 1

} else return X

(4)Step4: Repeat Step 1-3 until get the desired number of sample.

```{r HW6_4, warning=FALSE}
rffunc3 <- function(n, theta, beta_par) {
  alpha1 <- beta(theta, 1)
  alpha2 <- beta(1, beta_par) * sqrt(3)
  p1 <- alpha1/(alpha1+alpha2)
  p2 <- alpha2/(alpha1+alpha2)
  y <- rep(0, n)
  count <- 0
  while (count < n) {
    u <- runif(1, 0, 1)
    k <- sample(c(1, 2), size = 1, replace = 1, prob = c(p1, p2))
    if (k == 1) {
      x <- rbeta(1, shape1 = theta, shape2 = 1)
      if (u <= 1/(1+x^2)) {
        count <- count + 1
        y[count] <- x
      }
    } else {
     x <- rbeta(1, shape1 = 1, shape2 = beta_par)
     if (u <= sqrt((2+x^2)/3)) {
       count <- count + 1
       y[count] <- x
     }
    }
  }
  return(y)
}

X <- rffunc3(10000, theta = 2, beta_par = 1)
hist(X, nclass = 20, probability = TRUE)
points(density(X), type = "l", col = "red")
```
